---
title: "Data Analysis of Human Activity Recognition using Smartphone Dataset"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
    ```{r set-options, echo=FALSE, cache=FALSE}
options(width = 100)
```


## Data Set Description 

The present analysis is based on a data set coming from the reference: 
Davide Anguita, Alessandro Ghio, Luca Oneto, Xavier Parra and Jorge L. Reyes-Ortiz. Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine. International Workshop of Ambient Assisted Living (IWAAL 2012). Vitoria-Gasteiz, Spain. Dec 2012. 

The direct link to the variables description is in:
[links](http://archive.ics.uci.edu/ml/machine-learning-databases/00240/UCI%20HAR%20Dataset.names)

the data are downloaded from :
https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

First read data, labels and Subjects ID for train 

```{r include = TRUE}
library(dplyr)
dat1 <- read.table("./UCI HAR Dataset/train/X_train.txt")
labels1 <- read.table("./UCI HAR Dataset/train/Y_train.txt")
subject1 <- read.table("./UCI HAR Dataset/train/subject_train.txt")
dat1 <- cbind(subject1,dat1,labels1)
```
Then read data, labels and Subjects ID for test 
```{r include = TRUE}
dat2 <- read.table("./UCI HAR Dataset/test/X_test.txt")
labels2 <- read.table("./UCI HAR Dataset/test/Y_test.txt")
subject2 <- read.table("./UCI HAR Dataset/test/subject_test.txt")
dat2 <- cbind(subject2,dat2,labels2)
```
check if they really have the same number of features 
```{r include = TRUE}
dim(dat1)
dim(dat2)
```

read features names and activity labels
```{r include = TRUE}
feat <- read.table("./UCI HAR Dataset/features.txt")
features_names <- as.character(feat$V2)
class_labels <- read.table("./UCI HAR Dataset/activity_labels.txt")
class_labels <- class_labels$V2
```
check if the labels are factor and which levels they have
```{r include = TRUE}
levels(class_labels)
```

Combine train and test dataset

```{r include = TRUE}
dat <- rbind(dat1,dat2)
dim(dat)
```

add columns names + names of first and last one
```{r include = TRUE}
names(dat) <- c("subject_ID",features_names, "labelType")
```

remove duplicated column names
```{r include = TRUE}
dat <- dat[,!duplicated(names(dat))]
dim(dat)
```

select mean and std
```{r include = TRUE}
dat_small <- select(dat,c(ID = contains("subject_ID"),contains("mean"),contains("std"), activity_type = contains("labelType")))
dim(dat_small)
```

activity_type must be a factor
```{r include = TRUE}
dat_small$activity_type <- as.factor(dat_small$activity_type)
levels(dat_small$activity_type)
```

check level names
```{r include = TRUE}
levels(dat_small$activity_type) <- levels(class_labels)
# check if takes new levels
levels(dat_small$activity_type)
```

```{r include = TRUE}
# produce final averages for subjects and activities
dat_final <- dat_small %>% group_by(activity_type, ID) %>% summarize_all(mean)
```

```{r include = TRUE}
write.table(dat_final, file = "final.txt",  row.name = FALSE)
```

variables descriptions: 
the mean and sdt for the following variables are calculated for each type of activity and each subject ID: 
```{r include = TRUE}
colnames(dat_final)
```
